Slitasje og kjørespor

### 2023-01-19 DRAFT_slitasjeOgKjørespor.Rmd
feat: Check correlation between HIF and the indicator

```{r setup, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(units)
library(eaTools)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

@@

The figure above shows that Nord-Norge for example, has a lot of relatively untouched areas, and that Østlandet has the highest proportion of impacted areas.
This is expected. 
<!-- It also shows that the second HIF class is perhaps to narrowly defined. This is suggested as a change in another PR: -->

@@

```{r, eval=F}
slitasje_data_extrapolated <- ea_spread(indicator_data = slitasje_data,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 20)
saveRDS(slitasje_data_extrapolated, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated.rds")
```

```{r, echo=F}
# silently read cached file
slitasje_data_extrapolated <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated.rds")
```

```{r}
unique(slitasje_data_extrapolated$meanIndicatorValue)
```
<!-- slitasje_data_extrapolated contains 32402 rows, but I expected 5. -->

<!-- Like this: -->
```{r, eval=F}
slitasje_data_extrapolated2 <- ea_spread(indicator_data = slitasje_data_extrapolated,
                         indicator = meanIndicatorValue,
                         regions = regions,
                         groups = region,
                         threshold = 20)
saveRDS(slitasje_data_extrapolated2, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated2.rds")
```
```{r, echo=F}
# silently read cached file
slitasje_data_extrapolated2 <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated2.rds")
```

>>>>

```{r setup, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(units)
library(eaTools)
library(tidyverse)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

@@

The figure above shows that Nord-Norge for example, has a lot of relatively untouched areas, and that Østlandet has the highest proportion of impacted areas.
This is expected.

@@

```{r, eval=F}
slitasje_data_extrapolated <- ea_spread(indicator_data = slitasje_data,
                         indicator = i_2020,
                         regions = HIA_regions,
                         groups = region_HIF,
                         threshold = 20)
#saveRDS(slitasje_data_extrapolated, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated.rds")
```

Let's split the ID column to get the HIF
```{r}
slitasje_data_extrapolated <- slitasje_data_extrapolated %>%
  tidyr::separate(ID, into = c("region", "HIF"),
                  sep = " ")
slitasje_data_extrapolated$HIF <- as.numeric(slitasje_data_extrapolated$HIF)
```

Now let's check the correlation between the HIF and the indicator values. 
```{rslitasje-box, fig.cap="The relationship between human impact factor (a categorical version of the Norwegian Infrastructure Index) and the indicator avleus for slitasje."}
ggplot(slitasje_data_extrapolated, aes(x = factor(HIF), y = meanIndicatorValue))+
  geom_boxplot(fill = "grey",
               size=1.5)+
  theme_bw(base_size = 12)+
  labs(x = "Human Impact Factor", y = "Slitasje indicator value")
```

```{r}
cor(slitasje_data_extrapolated$HIF, slitasje_data_extrapolated$meanIndicatorValue)
```

This relationship is quite good, and I think it justified using HIF as homogeneous areas to extrapolate indicator values onto.

Split one more time, into regions:
```{r, eval=F}
slitasje_data_extrapolated2 <- ea_spread(indicator_data = slitasje_data_extrapolated,
                         indicator = meanIndicatorValue,
                         regions = regions,
                         groups = region,
                         threshold = 20)
saveRDS(slitasje_data_extrapolated2, "P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated2.rds")
```

```{r, echo=F}
# silently read cached file
slitasje_data_extrapolated2 <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/cache/slitasje_data_extrapolated2.rds")
```
